Time stretching of the GeV emission of GRBs: Fermi LAT data vs geometrical model

### 2014-12-12 Paper/gammaRays.tex
Added zmin value to the text

\begin{abstract}
	It is known that the high energy $(> \unit[100]{MeV})$
        emission of gamma-ray bursts is delayed with respect to the
        low energy emission.  However, the dependence of light curves
        on energy has not been studied for the high energy bands.  In
        this paper we consider the bursts observed by Fermi-LAT from
        2008 August 4 to 2011 August 1, for which at least $10$
        photons were observed with the energy greater than
        $\unit[1]{GeV}$. These include $4$ bursts: GRB 080916C, GRB
        090510, GRB 090902B, and GRB 090926A.  We use the
        Kolmogorov-Smirnov test to compare the light curves in the two
        bands, $\unit[100]{MeV} < E < \unit[1]{GeV}$ and
        $E > \unit[1]{GeV}$. For GRB 080916C and GRB
        090510 the light curves in the two bands are statistically
        compatible. However, for GRB 090926A, the higher-energy light
        curve is stretched compared to the lower-energy one with a
        statistical significance of $3.3 \sigma$ and, for GRB 090902B,
        on the contrary, the lower-energy curve is stretched with $2.3
        \sigma$ significance. We argue that the observed diversity of
        stretching factors may be explained in a simple geometrical
        model. The model assumes that the jet opening angle depends on
        the emission energy in a way that the most energetic photons
        are radiated near the axis of the jet. All the bursts are
        considered equivalent in their rest frames and the observed
        light curves differ only due to different redshifts and view
        directions. The model conforms to the total burst energy
        constraint and matches the Fermi LAT observations of the
        fraction of GRBs visible in $\unit[100]{MeV} < E <
        \unit[1]{GeV}$ band, which may be observed at higher
        energies. The model predicts the distribution of observable
        stretching factors, which may be tested in the future
        data. Finally, we propose a method to estimate observer's
        off-axis angle based on the stretching factor and the fraction of the
        high-energy photons. The code for modeling is open source and
        is publicly available on GitHub
        (\url{https://github.com/maxitg/GammaRays}).
\end{abstract}

@@

We start with selecting a range for redshifts. Ideally, one would
expect that this range starts with $z=0$. Then, however, there will be
visible jets for all possible off-axis angles, which will make our
angles range too large. Note also, that since the bursts count
increases with redshift as $z^3$, the probability to observe a
low-redshift burst is very small. So instead of selecting the smallest
redshift to be $0$, we select it to be a small number $z_\text{min}$.

>>>>

\begin{abstract}
	It is known that the high energy $(> \unit[100]{MeV})$
        emission of gamma-ray bursts is delayed with respect to the
        low energy emission.  However, the dependence of light curves
        on energy has not been studied for the high energy bands.  In
        this paper we consider the bursts observed by Fermi LAT from
        2008 August 4 to 2011 August 1, for which at least $10$
        photons were observed with the energy greater than
        $\unit[1]{GeV}$. These include $4$ bursts: GRB 080916C, GRB
        090510, GRB 090902B, and GRB 090926A.  We use the
        Kolmogorov-Smirnov test to compare the light curves in the two
        bands, $\unit[100]{MeV} < E < \unit[1]{GeV}$ and
        $E > \unit[1]{GeV}$. For GRB 080916C and GRB
        090510 the light curves in the two bands are statistically
        compatible. However, for GRB 090926A, the higher-energy light
        curve is stretched compared to the lower-energy one with a
        statistical significance of $3.3 \sigma$ and, for GRB 090902B,
        on the contrary, the lower-energy curve is stretched with $2.3
        \sigma$ significance. We argue that the observed diversity of
        stretching factors may be explained in a simple geometrical
        model. The model assumes that the jet opening angle depends on
        the emission energy in a way that the most energetic photons
        are radiated near the axis of the jet. All the bursts are
        considered equivalent in their rest frames and the observed
        light curves differ only due to different redshifts and view
        directions. The model conforms to the total burst energy
        constraint and matches the Fermi LAT observations of the
        fraction of GRBs visible in $\unit[100]{MeV} < E <
        \unit[1]{GeV}$ band, which may be observed at higher
        energies. The model predicts the distribution of observable
        stretching factors, which may be tested in the future
        data. Finally, we propose a method to estimate observer's
        off-axis angle based on the stretching factor and the fraction of the
        high-energy photons. The code for modeling is open source and
        is publicly available on GitHub
        (\url{https://github.com/maxitg/GammaRays}).
\end{abstract}

@@

We start with selecting a range for redshifts. Ideally, one would
expect that this range starts with $z=0$. Then, however, there will be
visible jets for all possible off-axis angles, which will make our
angles range too large. Note also, that since the bursts count
increases with redshift as $z^3$, the probability to observe a
low-redshift burst is very small. So instead of selecting the smallest
redshift to be $0$, we select it to be a small number $z_\text{min} = 0.1$.
